### 2025-04-07 Paper.tex
submission paper

\documentclass[a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\usepackage{vmargin}
\usepackage{multirow}	
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{comment}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{color}
\usepackage{framed}
\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{tikz}
\usepackage{float}
\usepackage{subcaption}
\usepackage{forest,calc}
\usepackage{tikz-qtree}
\usepackage{color}
\usepackage{colortbl}
\usepackage{capt-of}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlength{\headheight}{24pt}
\setlength{\headsep}{24pt}

\frenchspacing
\pagestyle{fancy}
\sloppy

\markright{headline}

% Kopfzeile
\lhead{\begin{tabular}{l}
\textbf{Software Project: Singing Voice Synthesis with DiffSinger}  \\
\subtitle \\
\end{tabular}}

\rhead{\begin{tabular}{rc} 
\StudNameOne\\
\StudNameTwo\\
\StudNameThree\\
\end{tabular}}

\author{\StudNumeOne}

\newcommand{\subtitle}{\textbf{Exercise 01}}
\newcommand{\StudNameOne}{\textbf{David Matkovic} - }
\newcommand{\StudNameTwo}{\textbf{Jona Hoppe} - 7034173}
\newcommand{\StudNameThree}{\textbf{Hanna Helbig} - 7008726}
\newcommand{\StudNameThree}{\textbf{Maximilian Schmidt} - }

\begin{document}

%==============================================
%================ Introduction =================
%==============================================
\section*{Introduction}

%==============================================
%================ Preprocessing ================
%==============================================
\section*{Preprocessing}

%==============================================
%================ Training =====================
%==============================================
\section*{Training}

%==============================================
%================ Evaluation ===================
%==============================================
\section*{Evaluation}

%==============================================
%============= Mixing and Mastering ============
%==============================================
\section*{Mixing and Mastering}

%==============================================
%================ Conclusion ===================
%==============================================
\section*{Conclusion}

\end{document}